How does color influence our political conversations?

### 2023-06-21 chapter_3/src/slides/slides.qmd
some more iterations of slides for ch3

---
title: "How does color influence our political conversations?"
author:
  - name: Damon C. Roberts
    title: PhD Candidate
    affiliations:
      - id: CU
        name: University of Colorado Boulder
        department: Political Science
        address: 333 UCB
        city: Boulder
        region: CO 
        postal-code: 80309-0333
    attributes:
      corresponding: true
format:
  revealjs: 
    embed-resources: true
    theme: [dark, ../../assets/slides/custom.scss]
    transition: slide
    scrollable: true
    incremental: true
    slide-number: true
    toc: true
    toc-depth: 1
    slide-level: 3
    bibliography: ../../../assets/references.bib
    footer: |
        <span class="faux-block">{{< fa brands creative-commons >}} 2021-2023 Damon Charles Roberts</span>
        <span class="faux-block"><br>All content licensed under {{< fa brands creative-commons >}} {{< fa brands creative-commons-by >}} {{< fa brands creative-commons-sa >}} [Creative Commons CC BY-SA 4.0](https://creativecommons.org/licenses/by-sa/4.0/)</span>
    self-contained: true
    output-file: color_and_conversation.html
execute:
  echo: false
---

@@

- Chapter from a larger book project
- Book examines
  - does color convey political information?
    - particularly partisan affiliation

@@

@@

@@

@@

@@

@@

@@

- Valance: a positive or negative response to an object.
- Affect (or affective state): feeling positive or negative in response to an object.
  - Pre-conscious response
- Emotion: Much more complex labels assigned to affective responses.
  - Often requires some amount of conscious processing to assign an appropriate label.
  - Slower
- Attitudes are an expression of valanced responses to some object.

## Learning from our psychology and neuroscience friends 😊

- There are a lot of useful things that we can borrow from social and affective neuroscientists and social psychologists

@@

1. Memory reflects stored prior experiences with various objects
  - These memories are not just factual recollections of things
    - e.g., the balloon was red.
  - Memories often encode your affective responses to that object.
2. Relatedly, memories are multi-dimensional
3. How we form an attitude with new information (or a combination of it) is by accessing memories that share features in common with the object.
4. Memories are a complex web of prior experiences that have a whole lot of strands connecting them.
5. Social groups and social hiearchies are a dimension that we pay attention to and encode along with a memory.

@@

@@

:::{#refs}
:::

>>>>

---
title: "How does color influence our political conversations?"
author:
  - name: Damon C. Roberts [{{< fa brands orcid size=xs >}}](https://orcid.org/0000-0002-4360-3675)
format:
  revealjs: 
    embed-resources: true
    theme: [dark, ../../assets/slides/custom.scss]
    transition: slide
    scrollable: false
    incremental: true
    slide-number: true
    toc: true
    toc-depth: 1
    slide-level: 3
    bibliography: ../../../assets/references.bib
    footer: |
        <span class="faux-block">{{< fa brands creative-commons >}} 2021-2023 Damon Charles Roberts</span>
        <span class="faux-block"><br>All content licensed under {{< fa brands creative-commons >}} {{< fa brands creative-commons-by >}} {{< fa brands creative-commons-sa >}} [Creative Commons CC BY-SA 4.0](https://creativecommons.org/licenses/by-sa/4.0/)</span>
    self-contained: true
    output-file: color_and_conversation.html
execute:
  echo: false
---

@@

::: footer
:::
---

::: {.notes}
- This is a chapter from a larger book project.
- The book addresses the question: 
  - does color convey political information? 
  - does it have pre-conscious effects on information that comes later?
:::

::: footer
:::

@@

::: {.notes}
- Ran a study on prolific
- Counterbalanced questions of attention, age, gender, race, colorblindness, partisan identity, etc.
- Ran three trials that used mousetracking software
- See one of three yard signs in each trial
  - Red, blue, or white yard sign
:::

::: footer
:::
---

@@

::: {.notes}
- Average predicted probabilities from two ordered logistic regressions
- Outcome is the respondent's belief that the candidate was either a Republican, Independent, or a Democrat.
- Read each figure by comparing the same colored bars. 
- Candidates using red yard signs are more likely to be seen as Republican
- Candidates using blue yard signs are more likely to be seen as Democrat
:::

::: footer
:::
---

@@

::: {.notes}
- The other two trials used less of the primary color and brought in more red or blue
- Do this as a test of the strength of these effects
- See that the results (kind of) weaken.
  - See that these effects weaken for the blue treatment. Much less change for the red treatment. Likely American flag.
:::

::: footer
:::
---

@@

::: {.notes}
- After viewing the three trials, participants are asked which of the three yard signs they prefer.
- Make a dichotomous variable of whether or not they like the first trial's yard sign.
- run logistic regressions to see whether votes are explained by the stimuli moderated by partisan identification
- See that there is a low predicted probability of strong Democrats voting for a candidate with a red yard sign.
  - See that the predicted probability of strong Republicans voting for a candidate with a red yard sign is high.
- See that there is a low predicted probability of strong republicans voting for a candidate with a blue yard sign.
  - See that the predicted probability of a strong Democrat voting for a candidate with a blue yard sign is high.
:::

::: footer
:::

@@

::: {.notes}
- Color seems to hold information about partisan affilations
  - Red = Republican
  - Blue = Democrat
- Unclear what sorts of effects this will have on attitudes and behaviors, though.
:::

::: footer
:::

@@

::: footer
:::

@@

- Valance: a positive or negative response to an object.
- Affect (or affective state): feeling positive or negative in response to an object.
  - Pre-conscious response
- Emotion: Much more complex labels assigned to affective responses.
  - Often requires some amount of conscious processing to assign an appropriate label.
  - Slower
- Attitudes are an expression of valanced responses to some object [@fazio_2007_sc].

::: footer
:::

## A cognitive architecture

::: footer
:::

@@

1. Memories are stored prior experiences [@cunningham_2021_yup].
  - often encode affective responses
2. Memories are multi-dimensional [@kahana_et-al_2022_ohhum].
3. Sensory and affective information is often encoded in memories [@kensinger_fields_2022_ohhum].
4. Memories are a complex network [@kahana_et-al_2022_ohhum].
  - Interconnected by different bits of information
5. Social groups and hierarchies are reflected in this network  [@santavirta_et-al_2023_n].

::: footer
:::

@@

::: {.notes}
Scenario.
- At the airport and watching the news
- Someone sits next to you before boarding
- They are wearing a red hat
- You might have one of two reactions
- A small conversation starts up since they see you watching the news.
- You see its a MAGA hat but they tell you they are having a hard time squaring some of their other beliefs with the stances of the party on immigration.
- You may shift your initial impression a bit. Feel more positive or more negative. However, you haven't fully reversed course on that. You have learned something though.

Implications.
- Color has downstream effects on our attitudes towards an object related to politics.
- The degree to which color matters in politics is whether or not it is primed.
:::

::: footer
:::

# Proposed research design

## Study 1

+---------------------:+:-----------------:+
| 🚫 Prime, Red shirt  | Prime, Red shirt  |
+----------------------+-------------------+
| 🚫 Prime, Blue shirt | Prime, Blue shirt |
+----------------------+-------------------+

: $2 \times 2$ factorial design {#tbl-study-1}

::: footer
:::

## Study 2

+-------------------------:+:---------------------:+
| Red shirt, 🚫 congruent  | Red shirt, congruent  |
+--------------------------+-----------------------+
| Blue shirt, 🚫 congruent | Blue shirt, congruent |
+--------------------------+-----------------------+

: $2 \times 2$ factorial design {#tbl-study-2}

::: footer
:::

@@

::: {#refs}
:::

::: footer
:::
